Breareal

### 2022-08-05 indicators/breareal-workingFile.Rmd
breareal

## About the underlying data
The indicator uses the maps of glacier extent \(_breatlas_\) from [2018-2019](https://www.nve.no/vann-og-vassdrag/vannets-kretsloep/bre/publikasjoner-publications/breatlas-glacier-inventories/) and compares it to a similar map from 1947-1985. Both maps are owned by NVE (Norsk vassdrags og energidirektorat). The new map is based on Sentinel images, and the old map is a digitised version of the N50 maps, with some additional areal photo interpretation on top.

@@

## Ecosystem characteristic
The indicator was initially designated to the _Abiotiske forhold_ characteristic, but now it is considered under _Landskapsøkologiske mønstre_. The initial justificaton was that glacial melting would influence water availability downstream. Now we put more weight on glaciers as not just resevoirs of ice and water, but as a habitat, and the presence or abundance of a specific habitat can be though of as an aspect of landscape state.

@@

## Collinearities with other indicators
None that are known

Probable or empirically documented

## Reference state
The indicator uses the same reference state as all or most of the indicators developed in this and in similar projects, and is defined in Nybø et al (2017): _Fagsystem for fastsetting av god økologisk tilstand_. As glacial extent is primarily, or solely, dependent on climate, we are in a case where we have actual data from the climatic reference period which is the previous normal period 1961-1990.

## Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
The reference value is zero change in glacial extent over time. The theshold for defining good ecological condition as set at a 40% reduction in glacial extnet, as a defult solution. This normalization follows a linear model, although a sigmoidal weibull function might be more appropriate.

The minumum value is set to zero.

## Uncertainties
Indicator values should ideally come with uncertainties associated with them, but in this case the data is very fixed so we do not attempt to calculate any uncertainties for this indicator.

@@

## Inporting the data sets
There are four data sets:

@@

We are assuming all glaciers are found in the mountins, se we do not need to mask the glacial atlsa data using an exosystem delineation map.

@@

There are 6915 polygons. Let's calculate the area of the polygons and study the distribution of size classes. 
```{r breareal}
breatlas$area <- st_area(breatlas)
par(mfrow=c(1,2))
hist(breatlas$area)
plot(breatlas$area)
```
Most polygons are small, and then there are a couple of very big ones. The larges is 50 km2.

@@

```
The red areas in the map to the far right is where the ice has retreated for a selected glacier on the west coast. 
Now we need to devide the map into regions to then compare the area in n50 with the new glacial extent map _breatlas_.

@@

Let me plot the new, fixed map on top of the old to see if the match. I give the new valid map a blue transparent colour, and the old map a yellow and transparent colour.  
```{r}
plot(nor$geometry, xlim=c(5000, 10000),
          ylim=c(6880000, 6885000),
     axes=T)
    plot(breatlas2$geometry, add=T, col = scales::alpha("blue",0.5))
    plot(breatlas$geometry, add=T, col = scales::alpha("yellow",0.5))
```
This seems to have worked. 
Notice also how some of these larger polygons are somewhat arbitrarely divided into several adjoining polygons. This is only so for the new map, not the old n50 map.

@@

Here is one method for getting the glacial area for each region.
```{r intersection}
# This intersection operation is a bit heavy, so I do it once and store it locally so that I can import it again quickly. 
#brealtas_reg <- st_intersection(breatlas, reg)
#saveRDS(brealtas_reg, "../data/brealtas_reg_helperfile.rds")

@@

Checking what happens to polygons that span to regions. Central Norway in red and western Norway in green. 50% transparency means we should get a third colour where and if they overlap.
```{r}
plot(brealtas_reg$geometry[brealtas_reg$region=="Midt-Norge"], 
    col = scales::alpha("red",0.5), border=NA, axes=T, 
     ylim=c(6900000, 6905000),
     xlim=c(85000, 100000))
plot(brealtas_reg$geometry[brealtas_reg$region=="Vestlandet"], 
     col = scales::alpha("green",0.5), border=NA, add=T)
  plot(nor$geometry, add=T)
```
It worked fine.

@@

# plotting
barplot(
  bretabell,
  ylab="Breareal (km2)"
)
```
Southern Norway has <1km2 of glaciers in 2018-2019.

### Same for N50
```{r intersection2}
n50x <- st_make_valid(n50)
#n50_reg <- st_intersection(n50x, reg)
#saveRDS(n50_reg, "../data/n50_helperfile.rds")
n50_reg <- readRDS("data/n50_helperfile.rds")
```

```{r}
plot(n50_reg$geometry[n50_reg$region=="Midt-Norge"], 
    col = scales::alpha("red",0.5), border=NA, axes=T, 
     ylim=c(6900000, 6905000),
     xlim=c(85000, 100000))
plot(n50_reg$geometry[n50_reg$region=="Vestlandet"], 
     col = scales::alpha("green",0.5), border=NA, add=T)
  plot(nor$geometry, add=T)
```
Thats good. 
Now I will plot the reference values (n50) and the contemporary data together.
```{r plotting-ref-and-ind}
n50_reg$area_crop <- st_area(n50_reg)

@@

In the figure above, the glacial area in the reference periond is in light grey and todays area is in dark grey. The height of the light grey bar depicts the reductino in glacial area.

There was a little more glaicers in southern Norway in the reference period, but only 4 km2. The condition here will be very poor, but this is not what we want, because glacial area is not at all representative of mountianareas in souther Norway.

## Scaled indicator values
Making a table with the indicator values.
```{r skalert}
myTbl <- tibble(region           = names(bretabell),
                indikator        = bretabell,
                referanseverdi   = n50tabell)
myTbl$skalert_indikator <- myTbl$indikator/myTbl$referanseverdi
myTbl
```

@@

Plotting an interactiv map with scaled indicator values
```{r}
tmap_mode("view")
tm_shape(reg_clipped) + 
  tm_polygons(col="skalert_indikator", border.col = "white")+
  tm_text("skalert_indikator_r", size=1.5)+
  tm_shape(nor)+
  tm_polygons(alpha = 0,border.col = "black")
```

Det er 4 måter å kalkulere den nasjonale indikatorverdien på:

1) ta gjennomsnitt av de skalerte regionale indikatorverdiene og la alle regionene telle likt (regional skaleringsmetode)
```{r}
(alt1 <- mean(reg_clipped$skalert_indikator))
```

2) summer indikatorverdien for hele landet og del på den summerte referanseverdien for hele landet (nasjonal metode)
```{r}
(alt2 <- sum(bretabell)/sum(n50tabell))
```

3) summere skalerte regionale indikatorverdier etter å ha gitt de ulike vekt grunnet totalarealet til regionen (arealveid metode)

```{r}
reg_clipped$areal <- st_area(reg_clipped$geometry)/1000000

totalArea <- sum(reg_clipped$areal)

reg_clipped$skalert_indikator_v <- reg_clipped$skalert_indikator*
  reg_clipped$areal

sumSkalerteIndikatorer <- sum(reg_clipped$skalert_indikator_v)

(alt3 <- sumSkalerteIndikatorer/totalArea)
```
4) summere skalerte regionale indikatorverdier etter å ha gitt de ulike vekt grunnet fjellareal (fjellarealveid metode)

For dette trenger jeg [fjellareal per region](fjellareal-per-region.html):
```{r}
(fjellareal <- readRDS("../data/fjellareal.rds"))
```

```{r}
reg_clipped$fjellareal <- fjellareal$Fjellareal[match(reg_clipped$region, fjellareal$Region)]
reg_clipped$skalert_indikator_v2 <- reg_clipped$skalert_indikator*
  reg_clipped$fjellareal
(alt4 <- sum(reg_clipped$skalert_indikator_v2)/sum(reg_clipped$fjellareal))
```
Her er alternativene oppsummert:

1. Indeksverdi `r round(alt1, 2)`. Aggregeringsmetoden er tvilsom.

2. Indeksverdi `r round(alt2, 2)`. Aggregeringsmetoden sier oss hvordan den TOTALE endringen i breareal har vært for Norge sett under ett.

3. Indeksverdi `r round(alt3, 2)`. Aggregeringsmetoden er igjen tvilsom.

4. Indeksverdi `r round(alt4, 2)`. Aggregeringsmetoden sier oss hvordan den GJENNOMSNITTLIGE endringen i breareal har vært for Norge sett under ett (jf metoden i NI).

Her kan det diskuteres hva som egentlig er best, men vi har valgt å gå for alt 4 for å være konsistent med SEEA EA bl.a. Indikatorverdien for Norge er da 0.62.

## Usikkerhet
NVE oppgir 3% usikkerhet ved metoden. Dette tolker jeg som at +- 3% utgjør 99.9%CI for estimatet (ca 2 SD).
```{r}
reg <- c("N", "C", "E", "W", "S")
myTbl$reg <- c("C", "N", "S", "W", "E")

breTbl <- data.frame(
  reg = rep(reg, each=10000),
  indikatorverdi = NA,
  refVerdi = NA,
  val = NA
)
breTbl2 <- breTbl

for(i in reg){
  #print(i)
  temp <- breTbl2[breTbl2$reg==i,]
  temp$indikatorverdi <- rnorm(10000, myTbl$indikator[myTbl$reg==i], myTbl$indikator[myTbl$reg==i]*0.015)
  temp$refVerdi <- rnorm(10000, myTbl$referanseverdi[myTbl$reg==i], myTbl$referanseverdi[myTbl$reg==i]*0.015)
  breTbl$indikatorverdi[breTbl$reg==i] <- temp$indikatorverdi
  breTbl$refVerdi[breTbl$reg==i] <- temp$refVerdi
}

# summary(breTbl$indikatorverdi) # no zeros
# summary(breTbl$refVerdi) # no zeros

breTbl$val <- breTbl$indikatorverdi/breTbl$refVerdi
breTbl$unscaledVal <- breTbl$refVerdi-breTbl$indikatorverdi
#breTbl <- dplyr::select(breTbl, -refVerdi)
breTbl$year <- 2019
```

Må jeg trunkere?
```{r}
summary(breTbl$val)
```
Nei, alle verdier er mellom 0 og 1.

Så må jeg finne den nasjonale tilstanden.

```{r}
breNorge <- data.frame(
  reg = rep("Norge", 10000),
  indikatorverdi = NA,
  refVerdi = NA,
  val = NA,
  unscaledVal = NA,
  year = 2019
)

# Legger inn vektene på nytt
wgt <- readRDS("../data/fjellareal.rds")
wgt$Fjellareal2 <- wgt$Fjellareal/max(wgt$Fjellareal)
wgt$reg <- c("N", "C", "E", "W", "S")

temp <- breTbl

temp2 <- c(
      sample(temp$val[temp$reg == "N"], wgt$Fjellareal2[wgt$reg == "N"]*1000, replace =T),
      sample(temp$val[temp$reg == "E"], wgt$Fjellareal2[wgt$reg == "E"]*1000, replace =T),
      sample(temp$val[temp$reg == "W"], wgt$Fjellareal2[wgt$reg == "W"]*1000, replace =T),
      sample(temp$val[temp$reg == "S"], wgt$Fjellareal2[wgt$reg == "S"]*1000, replace =T),
      sample(temp$val[temp$reg == "C"], wgt$Fjellareal2[wgt$reg == "C"]*1000, replace =T)
    )

for(i in 1:10000){

  breNorge$unscaledVal[i] <- 
    sum(
      sample(temp$unscaledVal[temp$reg == "N"],1),
      sample(temp$unscaledVal[temp$reg == "E"],1),
      sample(temp$unscaledVal[temp$reg == "W"],1),
      sample(temp$unscaledVal[temp$reg == "S"],1),
      sample(temp$unscaledVal[temp$reg == "C"],1))

  breNorge$refVerdi[i] <- 
    sum(
      sample(temp$refVerdi[temp$reg == "N"],1),
      sample(temp$refVerdi[temp$reg == "E"],1),
      sample(temp$refVerdi[temp$reg == "W"],1),
      sample(temp$refVerdi[temp$reg == "S"],1),
      sample(temp$refVerdi[temp$reg == "C"],1))

   breNorge$indikatorverdi[i] <- 
    sum(
      sample(temp$indikatorverdi[temp$reg == "N"],1),
      sample(temp$indikatorverdi[temp$reg == "E"],1),
      sample(temp$indikatorverdi[temp$reg == "W"],1),
      sample(temp$indikatorverdi[temp$reg == "S"],1),
      sample(temp$indikatorverdi[temp$reg == "C"],1))
}

breNorge$val <- sample(temp2, 10000, replace=T)

breTbl <- rbind(breTbl, breNorge)

```
# Eksport CSV
```{r, eval=F}
expdat <- dplyr::select(breTbl, reg, year, val)
write.csv(expdat, "../output/indicator_values/breareal.csv")
```

# Plotting
```{r}
eval(parse("indicator_plots2.R", encoding="UTF-8"))
```

```{r, message=F, eval=F}
png("../output/indicatorPlots/breareal.png", 
    units="in", width=5, height=7, res=300)

par(mfrow=c(1,1), mar=c(4.5,
                        5.5,
                        0,
                        2))

indicator_plot2(dataset = dplyr::select(breTbl, reg, year, val),
               yAxisTitle = "Breareal skalert mot referanseverdi",
               lowYlimit = 0,
               upperYlimit = 1,
               yStep = .2,
               minyear = 2018,
               maxyear = 2022,
               colours = c("#FFB25B", "#2DCCD3", "#004F71", "#7A9A01", "#93328E", "dark grey"),
               legendPosition = "right",
               legendInset = 0,
               move = 0.1,
               horizontal = F,
               legendTextSize = 1.25)
dev.off()
```

```{r}
png("../output/indicatorPlots/breareal_uskalert.png", 
    units="in", width=5, height=7, res=300)

par(mfrow=c(1,1), mar=c(4.5,
                        5.5,
                        2,
                        2))
indicator_plot2(dataset = dplyr::select(breTbl, reg, year, val = unscaledVal),
               yAxisTitle = expression(paste("Tapt breareal (km "^"-2"*")")),
               lowYlimit = 0,
               upperYlimit = 1100,
               yStep = 100,
               minyear = 2018,
               maxyear = 2022,
               colours = c("#FFB25B", "#2DCCD3", "#004F71", "#7A9A01", "#93328E", "dark grey"),
               legendPosition = "right",
               legendInset = 0,
               move = 0.1,
               horizontal = F,
               legendTextSize = 1.25)
dev.off()
```

```{r}
breTblTemp <- data.table::melt(breTbl,
                   id.vars = c("year", "reg"),
                   measure.vars = c("refVerdi", "indikatorverdi"),
                   variable.name = "year",
                   value.name = "val")
breTblTemp <- breTblTemp[,-1]

breTblTemp$year <- as.character(breTblTemp$year)
breTblTemp$year[breTblTemp$year == "refVerdi"] <- "1969"
breTblTemp$year[breTblTemp$year == "indikatorverdi"] <- "2019"
breTblTemp$year <- as.numeric(breTblTemp$year)

png("../output/indicatorPlots/breareal_uskalert_tidsserie.png", 
    units="in", width=5, height=7, res=300)

par(mfrow=c(1,1), mar=c(4.5,
                        5.5,
                        2,
                        2))
indicator_plot2(dataset = dplyr::select(breTblTemp, reg, year, val = val),
               yAxisTitle = expression(paste("Breareal (km "^"-2"*")")),
               lowYlimit = 0,
               upperYlimit = 4500,
               yStep = 1000,
               minyear = 1955,
               maxyear = 2022,
               colours = c("#FFB25B", "#2DCCD3", "#004F71", "#7A9A01", "#93328E", "dark grey"),
               legendPosition = "topright",
               legendInset = 0,
               move = 0.1,
               horizontal = F,
               legendTextSize = 1.25)
dev.off()
```

# Oppsummering

```{r, echo=F}

finalTbl <- aggregate(data=breTbl,
          val~year+reg,
          FUN= function(x) round(
            quantile(x, c(.025, .5, .975)), 2))

finalTbl <- do.call(data.frame, finalTbl)
names(finalTbl) <- c("year", "reg", "low", "med", "upp")

DT::datatable(
  finalTbl, 
  extensions = "FixedColumns",
  options = list(
    scrollX = TRUE,
    scrollY=T,
    pageLength = 10
  ))
```

```{r}
finalTbl2 <- aggregate(data=breTbl,
          indikatorverdi~year+reg,
          FUN= function(x) round(
            quantile(x, c(.025, .5, .975)), 2))

finalTbl2 <- do.call(data.frame, finalTbl2)
names(finalTbl2) <- c("year", "reg", "val_low", "val_med", "val_upp")

finalTbl3 <- aggregate(data=breTbl,
          refVerdi~year+reg,
          FUN= function(x) round(
            quantile(x, c(.025, .5, .975)), 2))

finalTbl3 <- do.call(data.frame, finalTbl3)
names(finalTbl3) <- c("year", "reg", "ref_low", "ref_med", "ref_upp")

finalTbl4 <- cbind(finalTbl2, finalTbl3[,3:5])
```

```{r, eval=F}
myReg <- c("E", "S", "W", "C", "N", "Norge")
labs = c("Østlandet","Sørlandet","Vestlandet","Midt-Norge","Nord-Norge","Norge")

bar <- ggplot(finalTbl)+
  geom_bar(aes(x=factor(reg, level = myReg), y=med), stat="identity", alpha=0.7)+
  geom_errorbar(aes(x=factor(reg, level = myReg), y=med, ymin = low, ymax=upp),
                width=0.5,
                size=1.2)+
  ylab("Breareal skalert mot referanseverdi")+
  xlab("")+
  scale_x_discrete(breaks = myReg,
                   labels= labs)+

  theme_bw(base_size = 20)+
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
  )

(barSTacked <- ggplot(finalTbl4)+
  geom_bar(aes(x=factor(reg, level = myReg), y=val_med), 
           stat="identity")+
  geom_errorbar(aes(x=factor(reg, level = myReg), 
                    y=val_med, ymin = val_low, ymax=val_upp),
                width=0.5,
                size=1.2)+

  geom_bar(aes(x=factor(reg, level = myReg), y=ref_med), 
           stat="identity", alpha=0.7)+
  geom_errorbar(aes(x=factor(reg, level = myReg), 
                    y=ref_med, ymin = ref_low, ymax=ref_upp),
                width=0.5,
                size=1.2)+

  ylab(expression(paste("Breareal (km "^"-2"*")")))+
  xlab("")+
  scale_x_discrete(breaks = myReg,
                   labels= labs)+
  theme_bw(base_size = 20)+
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
  ))

figure <- ggarrange(bar, barSTacked,
                    ncol = 2)

png("../output/indicatorPlots/breareal_barplot.png", 
    units="in", width=10, height=7, res=300)
figure
dev.off()
```

![](../output/indicatorPlots/breareal_barplot.png)

![](../output/indicatorPlots/breareal.png)

![](../output/indicatorPlots/breareal_uskalert.png)
![](../output/indicatorPlots/breareal_uskalert_tidsserie.png)

>>>>

## About the underlying data
The indicator uses the maps of glacier extent ( _breatlas_ ) from [2018-2019](https://www.nve.no/vann-og-vassdrag/vannets-kretsloep/bre/publikasjoner-publications/breatlas-glacier-inventories/) and compares it to a similar map from 1947-1985. Both maps are owned by NVE (Norsk vassdrags og energidirektorat). The new map is based on Sentinel images, and the old map is a digitised version of the N50 maps, with some additional areal photo interpretation on top.

@@

## Ecosystem characteristic
The indicator was initially assigned the _Abiotiske forhold_ characteristic, but now it is considered under _Landskapsøkologiske mønstre_. The initial justificaton was that glacial melting would influence water availability downstream. Now we put more weight on glaciers as not just resevoirs of ice and water, but as a habitat, and the presence or abundance of a specific habitat can be though of as an aspect of landscape state.

@@

## Collinearities with other indicators
None that are known.

## Reference state
The indicator uses the same reference state as all or most of the indicators developed in this and in similar projects, and is defined in [Nybø et al (2017): Fagsystem for fastsetting av god økologisk tilstand](https://www.regjeringen.no/no/dokumenter/fagsystem-for-fastsetting-av-god-okologisk-tilstand/id2558481/). As glacial extent is primarily, or solely, dependent on climate, we are in a case where we have actual data from the climatic reference period which is the previous normal period 1961-1990.

## Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
The reference value is zero change in glacial extent over time. The threshold for defining good ecological condition as set at a 40% reduction in glacial extent, as a default solution. This normalization follows a linear model, although a sigmoidal weibull function might be more appropriate.

The minimum value is set to zero.

## Uncertainties
Here we add a 3% error on each side of the estimates, same as what is reported for the glacier maps from NVE.

@@

## Importing the data sets
There are four data sets:

@@

We are assuming all glaciers are found in the mountins, se we do not need to mask the glacial atlas data using an ecosystem delineation map.

@@

There are 6915 polygons. Let's calculate the area of the polygons and study the distribution of size classes. 
```{r breareal}
breatlas$area <- st_area(breatlas)
par(mfrow=c(1,2))
hist(breatlas$area)
plot(breatlas$area)
```

Most polygons are small, and then there are a couple of very big ones. The largest polygon is 50 km2, but note that some polygons are somewhat arbitrarily split.

@@

The red areas in the map to the far right is where the ice has retreated for a selected glacier on the west coast. 
Now we need to devide the map into regions to then compare the area in n50 with the new glacial extent map _breatlas_.

@@

Let me plot the new, fixed map on top of the old to see if they match. I give the new valid map a blue transparent colour, and the old map a yellow and transparent colour.  
```{r}
plot(nor$geometry, xlim=c(5000, 10000),
          ylim=c(6880000, 6885000),
     axes=T)
    plot(breatlas2$geometry, add=T, col = scales::alpha("blue",0.5))
    plot(breatlas$geometry, add=T, col = scales::alpha("yellow",0.5))
```

This seems to have worked. 
Notice also how some of these larger polygons are somewhat arbitrarely divided into several adjoining polygons. This is only so for the new map, not the old n50 map.

@@

Here is one method for getting the glacial area for each region.
```{r intersection}
# This intersection operation is a bit heavy, so I do it once and store it locally so that I can import it again quickly. 
#brealtas_reg <- st_intersection(breatlas, reg)
#saveRDS(brealtas_reg, "indicators/data/brealtas_reg_helperfile.rds")

@@

Checking what happens to polygons that span to regions. Central Norway in red and western Norway in green. 50% transparency means we should get a third colour where and if they overlap.
```{r}
plot(brealtas_reg$geometry[brealtas_reg$region=="Midt-Norge"], 
    col = scales::alpha("red",0.5), border=NA, axes=T, 
     ylim=c(6900000, 6905000),
     xlim=c(85000, 100000))
plot(brealtas_reg$geometry[brealtas_reg$region=="Vestlandet"], 
     col = scales::alpha("green",0.5), border=NA, add=T)
  plot(nor$geometry, add=T)
```

It worked fine.

@@

# plotting
barplot(
  bretabell,
  ylab="Breareal (km2)"
)
```

Southern Norway has <1km2 of glaciers in 2018-2019.

### Same for N50
```{r intersection2}
n50x <- st_make_valid(n50)
#n50_reg <- st_intersection(n50x, reg)
#saveRDS(n50_reg, "indicators/data/n50_helperfile.rds")
n50_reg <- readRDS("data/n50_helperfile.rds")
```

```{r}
plot(n50_reg$geometry[n50_reg$region=="Midt-Norge"], 
    col = scales::alpha("red",0.5), border=NA, axes=T, 
     ylim=c(6900000, 6905000),
     xlim=c(85000, 100000))
plot(n50_reg$geometry[n50_reg$region=="Vestlandet"], 
     col = scales::alpha("green",0.5), border=NA, add=T)
  plot(nor$geometry, add=T)
```

Thats good. 
Now I will plot the reference values (n50) and the contemporary data together.
```{r plotting-ref-and-ind}
n50_reg$area_crop <- st_area(n50_reg)

@@

In the figure above, the glacial area in the reference periond is in light grey and todays area is in dark grey. The height of the light grey bar depicts the reduction in glacial area.

There was a little more glaicers in southern Norway in the reference period, but only 4 km2. The condition here will be very poor! I'm not entirely sure this is representative. The alternative is to set it to be NA, and I think I will do this - one could argue that glaciers are not as important of central to the alpine ecosystems in the south.

## Scaled indicator values
Making a table with the indicator values.
```{r skalert}
myTbl <- tibble(region           = names(bretabell),
                indikator        = bretabell,
                referanseverdi   = n50tabell)
myTbl$skalert_indikator <- myTbl$indikator/myTbl$referanseverdi

knitr::kable(myTbl)
```

@@

Plotting an interactive map with scaled indicator values
```{r}
tmap_mode("view")
tm_shape(reg_clipped) + 
  tm_polygons(col="skalert_indikator", 
              border.col = "white")+
  tm_text("skalert_indikator_r", size=1.5)+
  tm_shape(nor)+
  tm_polygons(alpha = 0,border.col = "black")
```

```{r}
# removing the value for souther norway
reg_clipped$skalert_indikator[reg_clipped$region=="Sørlandet"] <- NA
```

Calculating total change of glacier extent
```{r}
sum(bretabell)/sum(n50tabell)
```

We've lost 30% of the glacial area over this time.

## Uncertainty
NVE reports a 3% uncertainty with their method. This is interpret as if +- 3% represent a 99.9%CI for the estimate (ca 2 SD).
In the [original indicator](https://ninanor.github.io/IBECA/areal-av-isbreer.html) I re-sampled the indicator value 10000 times here, with a 0.015sd, to get a distribution of indicator values. This time I will settle for reporting the sd as a single value below.

# Prepare export
Now I can prepare the final exported data file. It will be a shape file with raw and scaled indicator values, reference reference values, and uncertainties.

The names of the columns are standardised (region, v_YYYY, sd_YYYY, i_YYYY, and reference).

```{r}
names(reg_clipped)
```

```{r}
# paste the raw variable estimates
reg_clipped$v_2019 <- myTbl$indikator[match(reg_clipped$region, myTbl$region)]

# paste the reference values
reg_clipped$reference <- myTbl$referanseverdi[match(reg_clipped$region, myTbl$region)]

# add 3% (one-sided) errors (SD=1.5% so that 1.95SD equals about 3%). 
reg_clipped$sd_2019 <- reg_clipped$v_2019*0.015

# subset the shape file to include only the columns of interest, and rename the column with the scaled indicator values, assigning the value to year 2019.
exp <- dplyr::select(reg_clipped,
                     region,
                     v_2019,
                     sd_2019,
                     reference,
                     i_2019 = skalert_indikator,
                     geometry)

## Eksport shape file (final product)

```{r, eval=F}
sf::st_write(exp, "indicatorMaps/breareal.shp")
```

```{r}
knitr::kable(exp[,1:5])
```
